(Deep) Induction Rules for GADTs

### 2021-12-14 induction-paper/cpp2022/cpp2022-paper-final.tex
minor tweaks to final version

\copyrightyear{2022}
\acmYear{2022}
\setcopyright{acmlicensed}\acmConference[CPP '22]{Proceedings of the 11th
ACM SIGPLAN International Conference on Certified Programs and
Proofs}{January 17--18, 2022}{Philadelphia, PA, USA}
\acmBooktitle{Proceedings of the 11th ACM SIGPLAN International Conference
on Certified Programs and Proofs (CPP '22), January 17--18, 2022,
Philadelphia, PA, USA}
\acmPrice{15.00}
\acmDOI{10.1145/3497775.3503680}
\acmISBN{978-1-4503-9182-5/22/01}

@@

\begin{table}[h!]
\begin{tabular}{|l|l|l|}
\hline
Data types & Discussed & Examples \\
          & in Sections &          \\
\hline\hline
ADTs & \ref{sec:intro} &
$\mathsf{List}$, $\mathsf{Rose}$ \\ 
Nested types & \ref{sec:ADTs-and-nesteds} & $\mathsf{PTree}$ \\  
Truly nested types & \ref{sec:ADTs-and-nesteds} & $\mathsf{Bush}$\\
GADTs & \ref{sec:GADTs} & $\mathsf{Eq}$, $\mathsf{Seq}$\\
Truly nested GADTs & \ref{sec:GADTs} and~\ref{sec:GADT-nested}&
$\mathsf{G}$ in \eqref{gadt-nested}\\ 
\hline
\end{tabular}
\caption{Data types in this paper}\label{fig:taxonomy}
\end{table}

@@

\begin{adjustbox}{varwidth=7.3in, max width=7.1in, margin=-0.0in 0in
      -0.2in 0in, center}

{\small
\[\begin{array}{l}
\mathsf{\forall (P : \forall (A : Set) \to (A \to Set) \to PTree\,A
  \to Set) \to \big( \forall (A : Set) (Q : A \to Set) (a : A) \to
  Q\,a \to P\,A\,Q\,(pleaf\, a) \big) \to} \\ \quad \mathsf{ \big(
  \forall (A : Set) (Q : A \to Set) (pp : PTree\,(A \times A)) \to
  P\,(A \times A)\,(Pair^{\wedge}\,A\,A\,Q\,Q)\,pp \to
  P\,A\,Q\,(pnode\,pp) \big) \to} \\ \quad \mathsf{\forall (A : Set) (Q
  : A \to Set) (p : PTree\,A) \to PTree^{\wedge}\,A\,Q\,p \to
  P\,A\,Q\,p }\\
 \\
\mathsf{\forall (P : \forall (A : Set) \to (A \to Set) \to Bush\, A \to Set)
\to \big( \forall (A : Set)\,(Q : A \to Set) \to P\,A\,Q\,bnil \big) \to} \\ 
\quad\mathsf{\big( \forall (A : Set) (Q : A \to Set) (a : A) (bb :
  Bush\,(Bush\,A)) \to Q\,a \to
  P\,(Bush\,A)\,(P\,A\,Q)\,bb \to P\,A\,Q\,(bcons\,a\,bb)
  \big) \to} \\ 
\quad\mathsf{\forall (A : Set) (Q : A \to Set) (b : Bush\,A) \to
  Bush^{\wedge}\,A\,Q\,b \to P\,A\,Q\,b } 
\end{array}\]}

\caption{Deep induction rules for perfect trees and
  bushes}\label{fig:ptree-and-bush}
\end{adjustbox}
\end{figure*}

@@

\begin{adjustbox}{varwidth=7.3in, max width=7.1in, margin=-0.0in 0in
      -0.2in 0in, center}

{\small
\[\begin{array}{l}
\mathsf{data\ LType : Set \to Set\ where}\\
\mathsf{\;\;bool :\, \forall \{A : Set\} \to \forall (B : Set) \to Equal\,A\,Bool
  \to LType\,A}\\ 
\mathsf{\;\;arr\;\;\; :\, \forall \{A : Set\} \to \forall (B\,C : Set) \to
  Equal\,A\,(B \to C) \to LType\,B \to LType\,C \to LType\,A}\\ 
  \mathsf{\;\;list\,\;\; :\,  \forall \{A : Set\} \to\forall (B : Set) \to
    Equal\,A\,(\List\,B) \to LType\,B \to LType\,A}\\
  \\
\mathsf{data\ LTerm : Set \to Set\ where}\\
\mathsf{\;\;var\,\,\,:\,  \forall \{A : Set\} \to String \to LType\,A \to
  LTerm\,A} \\  
\mathsf{\;\;abs\,\, :\,  \forall \{A : Set\} \to \forall (B\,C : Set) \to
  Equal\,A\,(B \to C) \to String \to LType\,B \to LTerm\,C \to
  LTerm\,A}\\ 
  \mathsf{\;\;app\; :\,  \forall \{A : Set\} \to \forall (B : Set) \to
    LTerm (B \to A) \to LTerm\,B \to LTerm\,A} \\ 
  \mathsf{\;\;list\;\,\, :\,  \forall \{A : Set\} \to \forall (B : Set) \to
    Equal\,A\,(List\,B) \to List\,(LTerm\,B) \to LTerm\,A} 
\end{array}\]}

\caption{The $\mathsf{LType}$ and $\mathsf{LTerm}$ data
  types}\label{fig:type-and-term}
\end{adjustbox}
\end{figure*}

@@

\begin{adjustbox}{varwidth=7.3in, max width=7.1in, margin=-0.0in 0in
      -0.2in 0in, center}

{\small
\[\begin{array}{lll}
\mathsf{Seq^{\wedge}\,A\,Q_A\,(const\,a)} & = & \mathsf{Q_A\,a}\\
\mathsf{Seq^{\wedge}\,A\,Q_A\,(pair\,B\,C\,e\,s_B\,s_C)}
&=&\mathsf{\exists [Q_B] \exists [Q_C]\, Equal^{\wedge}\,A\, (B
  \times C)\, Q_A\, (Pair^\wedge\,B\,C\, Q_B \, Q_C) \, e \times
  Seq^{\wedge}\,B\,Q_B\,s_B \times Seq^{\wedge}\,C\,Q_C\,s_C}\\
& & \\
\mathsf{dIndConst} & = & \mathsf{\lambda (P : \forall (A : Set) \to (A
  \to Set) \to Seq\,A \to Set) \to}\\
 & & \mathsf{\hspace*{0.15in}
\forall (A : Set) (Q_A : A \to Set) (a : A) \to Q_A\,a \to
P\,A\,Q_A\,(const\,a)}\\
& & \\
\mathsf{dIndPair} & = & \mathsf{\lambda (P : \forall (A : Set) \to (A
  \to Set) \to Seq\,A \to Set)} \to \\ 
 & & \quad \mathsf{\forall (A\,B\,C : Set) (Q_A : A \to Set) (Q_B : B
  \to Set) (Q_C : C \to Set)}\\
& & \mathsf{\hspace*{1.5in}(s_B : Seq\,B) (s_C : Seq\,C) (e :
  Equal\,A\,(B \times C)) \to} \\ 
& & \quad \mathsf{Equal^{\wedge} A\, (B \times C)\, Q_A\,
  (Pair^{\wedge}\,B\,C\,Q_B\,Q_C)\, e \to P\,B\,Q_B\,s_B \to}\\
& & \mathsf{\hspace*{1.5in}  P\,C\,Q_C\,s_C \to P\, A\, Q_A\, (
  pair\,B\,C\,e\,s_B\,s_C )} 
\end{array}\]

\vspace*{0.1in}

\[\begin{array}{l}
\hspace*{-2in}\mathsf{\forall (P : \forall (A : Set) \to (A \to Set) \to Seq\,A \to
  Set)} \mathsf{\to dIndConst\,P \to dIndPair\,P \to} \\ \hspace*{-2in}\quad
\mathsf{\forall (A : Set)(Q_A : A \to Set)(s_A : Seq\,A) \to
  Seq^{\wedge}\,A\,Q_A\,s_A \to P\,A\,Q_A\,s_A}
\end{array}\]}

\caption{Deep induction rule for $\mathsf{Seq}$}\label{fig:seq}
\end{adjustbox}
\end{figure*}

@@

%  \begin{adjustbox}{varwidth=6.8in, max width=6.6in, margin=-0.1in 0in
%      -0.2in 0in, fbox, center}

  \begin{adjustbox}{varwidth=7.3in, max width=7.1in, margin=-0.0in 0in
      -0.2in 0in, center}

  {\small
\[\begin{array}{lll}
\mathsf{LType^{\wedge}\,A\,Q_A\,(bool\,B\,e)} & = &\mathsf{\exists
  [Q_B]\, Equal^{\wedge}\, A\, B\, Q_A\, K^{Bool}_{\top} \,e}\\
\mathsf{LType^{\wedge}\,A\,Q_A\,(arr\, B\, C\, e\, T_B\, T_C)}
&=&\mathsf{\exists [Q_B] \,\exists [Q_c]\, Equal^{\wedge}\,A\,
  (B \to C)\, Q_A\, (Arr^{\wedge} \, B\, C\, Q_B \, Q_C) \, e \times
  \, LType^{\wedge}\,B\,Q_B\,T_B \times LType^{\wedge}\,C\,Q_C\,T_C}\\
\mathsf{LType^{\wedge}\,A\,Q_A\,(list\, B\, e\, T_B)} & = &
\mathsf{\exists [Q_B]\, Equal^{\wedge}\,A\, (List\, B)\, Q_A\,
  (List^{\wedge} \, B\, Q_B) \, e \times LType^{\wedge}\,B\,Q_B\,T_B}\\[1ex]
\mathsf{LTerm^{\wedge}\,A\,Q_A\,(var\,s\,T_A)} & = &
\mathsf{LType^{\wedge}\, A\, Q_A\, T_A}\\
\mathsf{LTerm^{\wedge}\,A\,Q_A\, (abs \,B \,C \,e \,s \,T_B \,t_C)} &
= & \mathsf{\exists [Q_B]\,\exists [Q_C]\, Equal^{\wedge} \, A\, (B \to
  C)\, Q_A\, (Arr^{\wedge} \, B\, C\, Q_B \, Q_C)\, e \times \,
  LType^{\wedge}\, B\, Q_B\, T_B \times \, LTerm^{\wedge}\, C\, Q_C\,
  t_C }\\
\mathsf{LTerm^{\wedge}\,A\,Q_A\, (app\, B\, t_{BA}\, t_B)} & = &
\mathsf{\exists [Q_B]\, LTerm^{\wedge}\, (B \to A)\, (Arr^{\wedge} \,
  B\, A\, Q_B \, Q_A)\, t_{BA} \times LTerm^{\wedge}\, B\, Q_B\,
  t_B}\\
\mathsf{LTerm^{\wedge}\,A\,Q_A\, (list\, B\, e\, ts)} & = &
\mathsf{\exists [Q_B]\, Equal^{\wedge} \, A\, (List\,B)\, Q_A\,
  (List^{\wedge} \, B\, Q_B) \, e \times List^{\wedge}\, (LTerm\,B) \,
  (LTerm^{\wedge} \, B\, Q_B) \, ts}
\end{array}   \]}

\caption{Predicate liftings for $\mathsf{LType}$ and
  $\mathsf{LTerm}$}\label{fig:liftings} \vspace*{0.1in} 
\end{adjustbox}
\end{figure*}

@@

\begin{adjustbox}{varwidth=7.3in, max width=7.1in, margin=-0.0in 0in
      -0.2in 0in, center}

{\small
\[\begin{array}{lll}
\mathsf{dIndVar} & = & \mathsf{\lambda (P : \forall (A : Set) \to (A \to Set) \to
  LTerm\,A  \to Set) \to}\\
& &   \quad\mathsf{\forall (A : Set) (Q_A : A \to Set) (s : String) (T_A :
  LType\, A) \to LType^{\wedge} \, A\, Q_A\, T_A \to P \, A\, Q_A\,
 (var \; s\, T_A)}\\[1ex]
\mathsf{dIndAbs} & = & \mathsf{\lambda (P : \forall (A : Set) \to (A \to Set) \to
  LTerm\,A \to Set) \to} \\ 
& &   \quad\mathsf{
  \forall (A\,B\,C: Set) (Q_A : A \to Set)  (Q_B : B \to Set) (Q_C : C
  \to Set) (e : Equal\, A\, (B \to C)) (s : String) \to } \\ 
& &   \quad\mathsf{(T_B : LType\, B) \to (t_C : LTerm\, C)
  \to Equal^{\wedge}\,A\,(B \to C)\, Q_A \, (Arr^{\wedge} \, B\, C\,
  Q_B \, Q_C) \, e \to  } \\
& &   \quad\mathsf{
  LType^{\wedge}\, B\, Q_B\, T_B
  \to P\, C\, Q_C\, t_C\, 
  \to P \, A\, Q_A\, (abs \,B \,C \, e \,s \,T_B \, t_C)}\\
 \\
\mathsf{dIndApp} & = &  \mathsf{\lambda (P : \forall (A : Set) \to (A \to Set) \to LTerm\,A
    \to Set)\to} \\ 
& &  \quad \mathsf{
  \forall (A \,B : Set) (Q_A : A \to Set)  (Q_B : B \to Set) 
   (t_{BA} : LTerm\, (B \to A)) (t_B : LTerm\, B) \to} \\
& &   \quad \mathsf{
  P\, (B \to A)\, (Arr^{\wedge} \, B\, A\, Q_B \, Q_A) \, t_{BA} \, 
  \to P\, B\, Q_B\, t_B\, 
  \to P \, A\, Q_A\, (app \,B \,t_{BA} \, t_B) }\\[1ex]
\mathsf{dIndList} & = &   \mathsf{\lambda (P : \forall (A : Set) \to (A \to Set) \to LTerm\,A
    \to Set) \to} \\ 
& &   \quad \mathsf{
  \forall (A \,B : Set) (Q_A : A \to Set)  (Q_B : B \to Set) 
    (e : Equal\, A\, (List\, B)) (ts : List\, (LTerm\, B)) \to} \\ 
& &   \quad \mathsf{
    Equal^{\wedge}\, A\, (List\,B)\, Q_A\, (List^{\wedge}\, B\, Q_B)\, e 
  \to List^{\wedge}\, (LTerm\,B) (P\, B\, Q_B)\, ts
  \to P \, A\, Q_A\, (list \,B \,e \, ts) }
\end{array}\]}

\caption{Induction hypotheses for $\mathsf{LTerm}$}\label{fig:ind-hyps-lterm}
\end{adjustbox}
\end{figure*}

    \begin{adjustbox}{varwidth=7.3in, max width=7.1in, margin=-0.0in 0in
      -0.2in 0in, center}

%  \begin{adjustbox}{varwidth=6.8in, max width=6.6in, margin=-0.1in 0in
%      -0.2in 0in, fbox, center} 
{\small
\[\begin{array}{lll}
\mathsf{dIndLTerm \, P\, cvar \, cabs\, capp\, clist \, A\, Q_A\,
  (var\;s\,T_A) \, liftA} & = & \mathsf{cvar \, A\, Q_A\, s\, T_A\,
  liftA}\\ 
\mathsf{dIndLTerm \, P\, cvar \, cabs\, capp\, clist \, A\, Q_A\,
  (abs \,B \,C \,e \,s \,T_B \, t_C) \, (Q_B , Q_C , liftE,
  lift_{T_B}, lift_{t_C})} & = & \mathsf{cabs\,A\,B\,C\, Q_A\,
  Q_B\, Q_C\, e\, s\, T_B\, t_C\, liftE\, lift_{T_B}\, p_C}\\
\mathsf{dIndLTerm \, P\, cvar \, cabs\, capp\, clist \, A\, Q_A\,
    (app \,B \,\,t_{BA} \, t_B)\, (Q_B , lift_{t_{BA}}, lift_{t_B})} &
= & \mathsf{capp\,A\,B\,Q_A\, Q_B\, t_{BA}\, t_B\, p_{BA} \, p_B}\\
  \mathsf{dIndLTerm \, P\, cvar \, cabs\, capp\, clist \, A\, Q_A\,
    (list \,B \,e \, ts) \, (Q_B , liftE', lift_{List})} & = & 
  \mathsf{clist \,A\,B\,Q_A\, Q_B\, e\, ts\, liftE'\, p_{List} }
\end{array}\]}
\mbox{where}
{\small
\[\begin{array}{lll}
\mathsf{p_C} & = & \mathsf{dIndLTerm\,P\,cvar\,cabs \,capp \,clist\,
  C\, Q_C\, t_C\, lift_{t_C} : P \, C\, Q_C \, t_C }\\
\mathsf{p_B} & = & \mathsf{dIndLTerm\,P\,cvar\,cabs \,capp \,clist\,
  B\, Q_B\, t_B\, lift_{t_B} : P \, B\, Q_B \, t_B }\\
\mathsf{p_{BA}} & = & \mathsf{dIndLTerm\,P\,cvar\,cabs \,capp
  \,clist\, (B \to A)\,(Arr^{\wedge} \, B\, A\, Q_B \, Q_A) \,
  t_{BA}\, lift_{t_{BA}} : P \, (B \to A)\, (Arr^{\wedge} \, B\, A\,
  Q_B \, Q_A) \, t_{BA}}\\ 
\mathsf{p_{List}} & = &\mathsf{liftListMap \, (LTerm\, B) \,
  (LTerm^{\wedge} \, B \, Q_B)\, (P\,B\,Q_B)\, p_{ts} \, ts\,
  lift_{List} : List^{\wedge}\, (LTerm\,B) \, (P\,B\,Q_B) \, ts}\\
\mathsf{p_{ts}} & = & \mathsf{dIndLTerm\, P\, cvar\, cabs\, capp\,
  clist\, B\, Q_B : PredMap\,(LTerm\,B) \,(LTerm^{\wedge}\, B\, Q_B)
  \, (P\,B\,Q_B)}
\end{array}\]}

\caption{$\mathsf{dIndLTerm}$}\label{fig:dindlterm} \vspace*{0.1in} 
\end{adjustbox}
\end{figure*}

To prove that this rule is sound we define a witness
$\mathsf{dIndLTerm}$ inhabiting the type in~\eqref{eq:ind-lam} as in
Figure~\ref{fig:dindlterm}. There,

>>>>

%%% The following is specific to CPP '22 and the paper
%%% '(Deep) Induction Rules for GADTs'
%%% by Patricia Johann and Enrico Ghiorzi.
%%%
\setcopyright{acmlicensed}
\acmPrice{15.00}
\acmDOI{10.1145/3497775.3503680}
\acmYear{2022}
\copyrightyear{2022}
\acmSubmissionID{poplws22cppmain-p17-p}
\acmISBN{978-1-4503-9182-5/22/01}
\acmConference[CPP '22]{Proceedings of the 11th ACM SIGPLAN International Conference on Certified Programs and Proofs}{January 17--18, 2022}{Philadelphia, PA, USA}
\acmBooktitle{Proceedings of the 11th ACM SIGPLAN International Conference on Certified Programs and Proofs (CPP '22), January 17--18, 2022, Philadelphia, PA, USA}

@@

\begin{table}[h!]
\caption{Data types in this paper}\label{fig:taxonomy}
\begin{tabular}{|l|l|l|}
\hline
Data types & Discussed & Examples \\
          & in Sections &          \\
\hline\hline
ADTs & \ref{sec:intro} &
$\mathsf{List}$, $\mathsf{Rose}$ \\ 
Nested types & \ref{sec:ADTs-and-nesteds} & $\mathsf{PTree}$ \\  
Truly nested types & \ref{sec:ADTs-and-nesteds} & $\mathsf{Bush}$\\
GADTs & \ref{sec:GADTs} & $\mathsf{Eq}$, $\mathsf{Seq}$\\
Truly nested GADTs & \ref{sec:GADTs} and~\ref{sec:GADT-nested}&
$\mathsf{G}$ in \eqref{gadt-nested}\\ 
\hline
\end{tabular}
\end{table}

@@

\begin{figure*}[t]
\begin{adjustbox}{varwidth=\linewidth, max width=\linewidth, center} 
{\small
\[\begin{array}{l}
\mathsf{\forall (P : \forall (A : Set) \to (A \to Set) \to PTree\,A
  \to Set) \to \big( \forall (A : Set) (Q : A \to Set) (a : A) \to
  Q\,a \to P\,A\,Q\,(pleaf\, a) \big) \to} \\ \quad \mathsf{ \big(
  \forall (A : Set) (Q : A \to Set) (pp : PTree\,(A \times A)) \to
  P\,(A \times A)\,(Pair^{\wedge}\,A\,A\,Q\,Q)\,pp \to
  P\,A\,Q\,(pnode\,pp) \big) \to} \\ \quad \mathsf{\forall (A : Set) (Q
  : A \to Set) (p : PTree\,A) \to PTree^{\wedge}\,A\,Q\,p \to
  P\,A\,Q\,p }\\
 \\
\mathsf{\forall (P : \forall (A : Set) \to (A \to Set) \to Bush\, A \to Set)
\to \big( \forall (A : Set)\,(Q : A \to Set) \to P\,A\,Q\,bnil \big) \to} \\ 
\quad\mathsf{\big( \forall (A : Set) (Q : A \to Set) (a : A) (bb :
  Bush\,(Bush\,A)) \to Q\,a \to
  P\,(Bush\,A)\,(P\,A\,Q)\,bb \to P\,A\,Q\,(bcons\,a\,bb)
  \big) \to} \\ 
\quad\mathsf{\forall (A : Set) (Q : A \to Set) (b : Bush\,A) \to
  Bush^{\wedge}\,A\,Q\,b \to P\,A\,Q\,b } 
\end{array}\]}
\end{adjustbox}
\caption{Deep induction rules for perfect trees and
  bushes}\label{fig:ptree-and-bush}
\end{figure*}

@@

\begin{figure*}[t]
\begin{adjustbox}{varwidth=\linewidth, max width=\linewidth, center} 
{\small
\[\begin{array}{l}
\mathsf{data\ LType : Set \to Set\ where}\\
\mathsf{\;\;bool :\, \forall \{A : Set\} \to \forall (B : Set) \to Equal\,A\,Bool
  \to LType\,A}\\ 
\mathsf{\;\;arr\;\;\; :\, \forall \{A : Set\} \to \forall (B\,C : Set) \to
  Equal\,A\,(B \to C) \to LType\,B \to LType\,C \to LType\,A}\\ 
  \mathsf{\;\;list\,\;\; :\,  \forall \{A : Set\} \to\forall (B : Set) \to
    Equal\,A\,(\List\,B) \to LType\,B \to LType\,A}\\
  \\
\mathsf{data\ LTerm : Set \to Set\ where}\\
\mathsf{\;\;var\,\,\,:\,  \forall \{A : Set\} \to String \to LType\,A \to
  LTerm\,A} \\  
\mathsf{\;\;abs\,\, :\,  \forall \{A : Set\} \to \forall (B\,C : Set) \to
  Equal\,A\,(B \to C) \to String \to LType\,B \to LTerm\,C \to
  LTerm\,A}\\ 
  \mathsf{\;\;app\; :\,  \forall \{A : Set\} \to \forall (B : Set) \to
    LTerm (B \to A) \to LTerm\,B \to LTerm\,A} \\ 
  \mathsf{\;\;list\;\,\, :\,  \forall \{A : Set\} \to \forall (B : Set) \to
    Equal\,A\,(List\,B) \to List\,(LTerm\,B) \to LTerm\,A} 
\end{array}\]}
\end{adjustbox}
\caption{The $\mathsf{LType}$ and $\mathsf{LTerm}$ data
  types}\label{fig:type-and-term}
\end{figure*}

@@

\begin{figure*}[t]
\begin{adjustbox}{varwidth=\linewidth, max width=\linewidth, center} 
{\small
\[\begin{array}{lll}
\mathsf{Seq^{\wedge}\,A\,Q_A\,(const\,a)} & = & \mathsf{Q_A\,a}\\
\mathsf{Seq^{\wedge}\,A\,Q_A\,(pair\,B\,C\,e\,s_B\,s_C)}
&=&\mathsf{\exists [Q_B] \exists [Q_C]\, Equal^{\wedge}\,A\, (B
  \times C)\, Q_A\, (Pair^\wedge\,B\,C\, Q_B \, Q_C) \, e \times
  Seq^{\wedge}\,B\,Q_B\,s_B \times Seq^{\wedge}\,C\,Q_C\,s_C}\\
& & \\
\mathsf{dIndConst} & = & \mathsf{\lambda (P : \forall (A : Set) \to (A
  \to Set) \to Seq\,A \to Set) \to}\\
 & & \mathsf{\hspace*{0.15in}
\forall (A : Set) (Q_A : A \to Set) (a : A) \to Q_A\,a \to
P\,A\,Q_A\,(const\,a)}\\
& & \\
\mathsf{dIndPair} & = & \mathsf{\lambda (P : \forall (A : Set) \to (A
  \to Set) \to Seq\,A \to Set)} \to \\ 
 & & \quad \mathsf{\forall (A\,B\,C : Set) (Q_A : A \to Set) (Q_B : B
  \to Set) (Q_C : C \to Set)}\\
& & \mathsf{\hspace*{1.5in}(s_B : Seq\,B) (s_C : Seq\,C) (e :
  Equal\,A\,(B \times C)) \to} \\ 
& & \quad \mathsf{Equal^{\wedge} A\, (B \times C)\, Q_A\,
  (Pair^{\wedge}\,B\,C\,Q_B\,Q_C)\, e \to P\,B\,Q_B\,s_B \to}\\
& & \mathsf{\hspace*{1.5in}  P\,C\,Q_C\,s_C \to P\, A\, Q_A\, (
  pair\,B\,C\,e\,s_B\,s_C )} 
\end{array}\]
\[\begin{array}{l}
\hspace*{-2in}\mathsf{\forall (P : \forall (A : Set) \to (A \to Set) \to Seq\,A \to
  Set)} \mathsf{\to dIndConst\,P \to dIndPair\,P \to} \\ \hspace*{-2in}\quad
\mathsf{\forall (A : Set)(Q_A : A \to Set)(s_A : Seq\,A) \to
  Seq^{\wedge}\,A\,Q_A\,s_A \to P\,A\,Q_A\,s_A}
\end{array}\]}
\end{adjustbox}
\caption{Deep induction rule for $\mathsf{Seq}$}\label{fig:seq}
\end{figure*}

@@

\begin{figure*}[t]
\begin{adjustbox}{varwidth=\linewidth, max width=\linewidth, center} 
  {\small
\[\begin{array}{lll}
\mathsf{LType^{\wedge}\,A\,Q_A\,(bool\,B\,e)} & = &\mathsf{\exists
  [Q_B]\, Equal^{\wedge}\, A\, B\, Q_A\, K^{Bool}_{\top} \,e}\\
\mathsf{LType^{\wedge}\,A\,Q_A\,(arr\, B\, C\, e\, T_B\, T_C)}
&=&\mathsf{\exists [Q_B] \,\exists [Q_c]\, Equal^{\wedge}\,A\,
  (B \to C)\, Q_A\, (Arr^{\wedge} \, B\, C\, Q_B \, Q_C) \, e \times
  \, LType^{\wedge}\,B\,Q_B\,T_B \times LType^{\wedge}\,C\,Q_C\,T_C}\\
\mathsf{LType^{\wedge}\,A\,Q_A\,(list\, B\, e\, T_B)} & = &
\mathsf{\exists [Q_B]\, Equal^{\wedge}\,A\, (List\, B)\, Q_A\,
  (List^{\wedge} \, B\, Q_B) \, e \times LType^{\wedge}\,B\,Q_B\,T_B}\\[1ex]
\mathsf{LTerm^{\wedge}\,A\,Q_A\,(var\,s\,T_A)} & = &
\mathsf{LType^{\wedge}\, A\, Q_A\, T_A}\\
\mathsf{LTerm^{\wedge}\,A\,Q_A\, (abs \,B \,C \,e \,s \,T_B \,t_C)} &
= & \mathsf{\exists [Q_B]\,\exists [Q_C]\, Equal^{\wedge} \, A\, (B \to
  C)\, Q_A\, (Arr^{\wedge} \, B\, C\, Q_B \, Q_C)\, e \times \,
  LType^{\wedge}\, B\, Q_B\, T_B \times \, LTerm^{\wedge}\, C\, Q_C\,
  t_C }\\
\mathsf{LTerm^{\wedge}\,A\,Q_A\, (app\, B\, t_{BA}\, t_B)} & = &
\mathsf{\exists [Q_B]\, LTerm^{\wedge}\, (B \to A)\, (Arr^{\wedge} \,
  B\, A\, Q_B \, Q_A)\, t_{BA} \times LTerm^{\wedge}\, B\, Q_B\,
  t_B}\\
\mathsf{LTerm^{\wedge}\,A\,Q_A\, (list\, B\, e\, ts)} & = &
\mathsf{\exists [Q_B]\, Equal^{\wedge} \, A\, (List\,B)\, Q_A\,
  (List^{\wedge} \, B\, Q_B) \, e \times List^{\wedge}\, (LTerm\,B) \,
  (LTerm^{\wedge} \, B\, Q_B) \, ts}
\end{array}   \]}
\end{adjustbox}
\caption{Predicate liftings for $\mathsf{LType}$ and
  $\mathsf{LTerm}$}\label{fig:liftings}
\end{figure*}

@@

\begin{adjustbox}{varwidth=\linewidth, max width=\linewidth, center} 
{\small
\[\begin{array}{lll}
\mathsf{dIndVar} & = & \mathsf{\lambda (P : \forall (A : Set) \to (A \to Set) \to
  LTerm\,A  \to Set) \to}\\
& &   \quad\mathsf{\forall (A : Set) (Q_A : A \to Set) (s : String) (T_A :
  LType\, A) \to LType^{\wedge} \, A\, Q_A\, T_A \to P \, A\, Q_A\,
 (var \; s\, T_A)}\\[1ex]
\mathsf{dIndAbs} & = & \mathsf{\lambda (P : \forall (A : Set) \to (A \to Set) \to
  LTerm\,A \to Set) \to} \\ 
& &   \quad\mathsf{
  \forall (A\,B\,C: Set) (Q_A : A \to Set)  (Q_B : B \to Set) (Q_C : C
  \to Set) (e : Equal\, A\, (B \to C)) (s : String) \to } \\ 
& &   \quad\mathsf{(T_B : LType\, B) \to (t_C : LTerm\, C)
  \to Equal^{\wedge}\,A\,(B \to C)\, Q_A \, (Arr^{\wedge} \, B\, C\,
  Q_B \, Q_C) \, e \to  } \\
& &   \quad\mathsf{
  LType^{\wedge}\, B\, Q_B\, T_B
  \to P\, C\, Q_C\, t_C\, 
  \to P \, A\, Q_A\, (abs \,B \,C \, e \,s \,T_B \, t_C)}\\
 \\
\mathsf{dIndApp} & = &  \mathsf{\lambda (P : \forall (A : Set) \to (A \to Set) \to LTerm\,A
    \to Set)\to} \\ 
& &  \quad \mathsf{
  \forall (A \,B : Set) (Q_A : A \to Set)  (Q_B : B \to Set) 
   (t_{BA} : LTerm\, (B \to A)) (t_B : LTerm\, B) \to} \\
& &   \quad \mathsf{
  P\, (B \to A)\, (Arr^{\wedge} \, B\, A\, Q_B \, Q_A) \, t_{BA} \, 
  \to P\, B\, Q_B\, t_B\, 
  \to P \, A\, Q_A\, (app \,B \,t_{BA} \, t_B) }\\[1ex]
\mathsf{dIndList} & = &   \mathsf{\lambda (P : \forall (A : Set) \to (A \to Set) \to LTerm\,A
    \to Set) \to} \\ 
& &   \quad \mathsf{
  \forall (A \,B : Set) (Q_A : A \to Set)  (Q_B : B \to Set) 
    (e : Equal\, A\, (List\, B)) (ts : List\, (LTerm\, B)) \to} \\ 
& &   \quad \mathsf{
    Equal^{\wedge}\, A\, (List\,B)\, Q_A\, (List^{\wedge}\, B\, Q_B)\, e 
  \to List^{\wedge}\, (LTerm\,B) (P\, B\, Q_B)\, ts
  \to P \, A\, Q_A\, (list \,B \,e \, ts) }
\end{array}\]}
\end{adjustbox}
\caption{Induction hypotheses for $\mathsf{LTerm}$}\label{fig:ind-hyps-lterm}
\end{figure*}

\begin{figure*}[t]
\begin{adjustbox}{varwidth=\linewidth, max width=\linewidth, center} 
%  \begin{adjustbox}{varwidth=6.8in, max width=6.6in, margin=-0.1in 0in
%      -0.2in 0in, fbox, center} 
{\small
\[\begin{array}{lll}
\mathsf{dIndLTerm \, P\, cvar \, cabs\, capp\, clist \, A\, Q_A\,
  (var\;s\,T_A) \, liftA} & = & \mathsf{cvar \, A\, Q_A\, s\, T_A\,
  liftA}\\ 
\mathsf{dIndLTerm \, P\, cvar \, cabs\, capp\, clist \, A\, Q_A\,
  (abs \,B \,C \,e \,s \,T_B \, t_C) \, (Q_B , Q_C , liftE,
  lift_{T_B}, lift_{t_C})} & = & \mathsf{cabs\,A\,B\,C\, Q_A\,
  Q_B\, Q_C\, e\, s\, T_B\, t_C\, liftE\, lift_{T_B}\, p_C}\\
\mathsf{dIndLTerm \, P\, cvar \, cabs\, capp\, clist \, A\, Q_A\,
    (app \,B \,\,t_{BA} \, t_B)\, (Q_B , lift_{t_{BA}}, lift_{t_B})} &
= & \mathsf{capp\,A\,B\,Q_A\, Q_B\, t_{BA}\, t_B\, p_{BA} \, p_B}\\
  \mathsf{dIndLTerm \, P\, cvar \, cabs\, capp\, clist \, A\, Q_A\,
    (list \,B \,e \, ts) \, (Q_B , liftE', lift_{List})} & = & 
  \mathsf{clist \,A\,B\,Q_A\, Q_B\, e\, ts\, liftE'\, p_{List} }
\end{array}\]}
\mbox{where}
{\small
\[\begin{array}{lll}
\mathsf{p_C} & = & \mathsf{dIndLTerm\,P\,cvar\,cabs \,capp \,clist\,
  C\, Q_C\, t_C\, lift_{t_C} : P \, C\, Q_C \, t_C }\\
\mathsf{p_B} & = & \mathsf{dIndLTerm\,P\,cvar\,cabs \,capp \,clist\,
  B\, Q_B\, t_B\, lift_{t_B} : P \, B\, Q_B \, t_B }\\
\mathsf{p_{BA}} & = & \mathsf{dIndLTerm\,P\,cvar\,cabs \,capp
  \,clist\, (B \to A)\,(Arr^{\wedge} \, B\, A\, Q_B \, Q_A) \,
  t_{BA}\, lift_{t_{BA}} : P \, (B \to A)\, (Arr^{\wedge} \, B\, A\,
  Q_B \, Q_A) \, t_{BA}}\\ 
\mathsf{p_{List}} & = &\mathsf{liftListMap \, (LTerm\, B) \,
  (LTerm^{\wedge} \, B \, Q_B)\, (P\,B\,Q_B)\, p_{ts} \, ts\,
  lift_{List} : List^{\wedge}\, (LTerm\,B) \, (P\,B\,Q_B) \, ts}\\
\mathsf{p_{ts}} & = & \mathsf{dIndLTerm\, P\, cvar\, cabs\, capp\,
  clist\, B\, Q_B : PredMap\,(LTerm\,B) \,(LTerm^{\wedge}\, B\, Q_B)
  \, (P\,B\,Q_B)}
\end{array}\]}
\end{adjustbox}
\caption{$\mathsf{dIndLTerm}$}\label{fig:dindlterm} \vspace*{0.1in} 
\end{figure*}

To prove this rule sound we define a witness
$\mathsf{dIndLTerm}$ inhabiting the type in~\eqref{eq:ind-lam} as in
Figure~\ref{fig:dindlterm}. There,
